Lab 10: Poisson and Negative Binomial Models: Answers

### 2023-03-29 Lab 10 PoissonNBmodels/Lab10_PoissonNBmodels.Rmd
Add files via upload

---
title: "Lab 10: Poisson and Negative Binomial Models"
author: "Eric Parajon"
output: pdf_document
---

@@

Load necessary packages:
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, messsage = FALSE)
rm(list=ls())

#Load in packages
pacman::p_load(tidyverse,
               haven,#for importing data
               labelled,
               MASS,
               AER,
               foreign)
```

@@

Nonetheless, let's build our model. We're going to look at the number of militarized interstate wars based on 'trade network polarization' (lgtrdnpi2) and the number of democratic cliques (clqdemreg) in the system:

```{r}
m1 <- glm(nowars20 ~ lgtrdnpi2 + clqdemreg,
          data = count,
          family = "poisson")
summary(m1)
```
Note that this tells us "Dispersion parameter for poisson family taken to be 1."

@@

--- don't look ----
--- don't look ----
--- don't look ----
--- don't look ----
--- don't look ----
--- don't look ----
--- don't look ----
--- don't look ----
--- don't look ----
--- don't look ----
--- don't look ----
--- don't look ----
--- don't look ----
--- don't look ----

@@

m2<- glm.nb(nowars20 ~ lgtrdnpi2 + clqdemreg, 
            data = count)
summary(m2)

@@

```{r}
install.packages("countreg", repos="http://R-Forge.R-project.org")
library(countreg)
library(cowplot)

@@

ylims <- ylim(-5, 11)  # common scale for comparison
plot_grid(autoplot(root.pois) + ylims, autoplot(root.nb) + ylims, 
          ncol = 2, labels = c("Poisson", "Neg. Binom"))

```
What does this tell us? The rootogram reflects fitted counts. If a bar doesn't reach the zero line, the model is overpredicting a certain count bin. If it exceeds the zero line, it underpredicts. So, neither of these models are perfect. BUT - the poisson model seems to both over and underpredict a lot. The negative binomial model on the other hand definitely over predicts, but not as much.

@@

#What if we wanted a single point estimate of our predicted count of wars given our IVs?
mean(pc_sims)

#And the upper and lower boundaries of a confidence interval based on this simulation?
quantile(pc_sims, prob= 0.025) 
quantile(pc_sims, prob= 0.975) 
```

How do we predict counts over the range of a variable? Like we always have! Let's say we want to know what happens to the predicted number of counts as the average number of democratic cliques varies from its min to its max.
```{r}
temp_data <- count

@@

pc_dem_cl <- purrr::map_dfr(dem_cl, function(x){

@@

#we could do this in one step, but I like keeping all of the simulated counts so
#we can do whatever we want with the distribution of predicted counts
plot_data <- t(apply(pc_dem_cl, 1, quantile, probs = c(.1, .5, .9))) %>% as.data.frame()
plot_data$dem_cl <- dem_cl
```

@@

Below, try using ggplot to plot this information we just generated. Follow these guidelines:
- X axis: Average number of Democratic Cliques
- Y axis: Predicted count of Intl. Wars
- Include a geom_line() for Dem_Cliques
- Include a geom_ribbon() with CIs (80% confidence interval)
- Include a legend which specifies what the line is and what the CIs are, try putting that legend on the bottom

```{r}
#Now to plot: 
colnames(plot_data)<-c("low", "med", "high", "Dem_Cliques")

>>>>

---
title: 'Lab 10: Poisson and Negative Binomial Models'
author: "Eric Parajon"
output:
  html_document:
    df_print: paged
---

@@

Load necessary packages:
```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE, messsage = FALSE)
rm(list=ls())

#Load in packages
pacman::p_load(tidyverse,
               haven,#for importing data
               labelled,
               MASS,
               AER,
               foreign,
               countreg,
               cowplot)
```

@@

Nonetheless, let's build our model. We're going to look at the number of militarized interstate wars based on 'trade network polarization' (lgtrdnpi2) and the number of democratic cliques (clqdemreg) in the system: 
Since we're using count data what should the family argument be?

```{r}
m1 <- glm(nowars20 ~ lgtrdnpi2 + clqdemreg,
          data = count,
          family = "poisson")
summary(m1)

```
Note that this tells us "Dispersion parameter for poisson family taken to be 1."

@@

--- don't look\
--- don't look\
--- don't look\
--- don't look\
--- don't look\
--- don't look\
--- don't look\
--- don't look\
--- don't look\
--- don't look\
--- don't look\

@@

m2<- MASS::glm.nb(nowars20 ~ lgtrdnpi2 + clqdemreg, 
            data = count)
summary(m2)

@@

```{r}
#install.packages("countreg", repos="http://R-Forge.R-project.org")

library(countreg)
library(cowplot)

@@

ylims <- ylim(-5, 11)  # common scale for comparison
#Using cowplot to put plots together with labels
plot_grid(autoplot(root.pois) + ylims, autoplot(root.nb) + ylims, 
          ncol = 2, labels = c("Poisson", "Neg. Binom"))

```
What does this tell us?

The rootogram reflects fitted counts. If a bar doesn't reach the zero line, the model is overpredicting a certain count bin. If it exceeds the zero line, it underpredicts. So, neither of these models are perfect. BUT - the poisson model seems to both over and underpredict a lot. The negative binomial model on the other hand definitely over predicts, but not as much.

@@

#What if we wanted a single point estimate of our predicted count of wars given our IVs?
mean<-mean(pc_sims)

#And the upper and lower boundaries of a confidence interval based on this simulation?
lower<-quantile(pc_sims, prob= 0.025) 
upper<-quantile(pc_sims, prob= 0.975)

#putting it together
mapply(c,mean,lower,upper)

How do we predict counts over the range of a variable? Like we always have! Let's say we want to know what happens to the predicted number of counts as the average number of democratic cliques varies from its min to its max.
```{r}

#Re-naming count df to temp_data for use in this
temp_data <- count

@@

#Using map_dfr to directly binds the rows to df we're calling pc_dem_cl
pc_dem_cl <- purrr::map_dfr(dem_cl, function(x){

@@

#we could do this in one step, but I like keeping all of the simulated counts so
#we can do whatever we want with the distribution of predicted counts

plot_data <- t(apply(pc_dem_cl, 1, quantile, probs = c(.1, .5, .9))) %>% as.data.frame()
plot_data$dem_cl <- dem_cl

#So this give us the average point estimate (expected count), the low bound of a confidence interval, and the high bound at varying numbers of democratic cliques

# Exercise

@@

Below, try using ggplot to plot this information we just generated. Follow these guidelines:\
- X axis: Average number of Democratic Cliques\
- Y axis: Predicted count of Intl. Wars\
- Include a geom_line() for Dem_Cliques\
- Include a geom_ribbon() with CIs (80% confidence interval)\
- Include a legend which specifies what the line is and what the CIs are, try putting that legend on the bottom

```{r, eval=FALSE}
#Now to plot the above data :

#Renaming columns
colnames(plot_data)<-c("low", "expected", "high", "Dem_Cliques")

### 2023-03-29 Lab 10 PoissonNBmodels/Lab10_PoissonNBmodels_AnswerKey.Rmd
Add files via upload

---
title: "Lab 10: Poisson and Negative Binomial Models: Answers"
author: "TA: Eric Parajon"
output: pdf_document
---

@@

Load necessary packages:
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, messsage = FALSE)
rm(list=ls()) 
#Load in packages
pacman::p_load(tidyverse,
               haven,#for importing data
               labelled,
               MASS,
               AER,
               foreign)
```

@@

```{r, message=F, warning=F, include=F}
count <- read_dta("Maoz.dta")
#count2 <-na.omit(count) # <- don't do this with your data unless you are absolutely
#sure that you want to drop all nas 
```

@@

```{r}
m1 <- glm(data = count, nowars20 ~ lgtrdnpi2 + clqdemreg, family="poisson")
summary(m1)
```
Note that this tells us "Dispersion parameter for poisson family taken to be 1."

@@

ggplot(data=over, aes(x=fits, y=V2))+
  geom_point()+
  geom_smooth(aes(x=fits, y=V2), method=lm, se=FALSE, color="blue")+ #line for mean of estimates against the variance of estimates
  geom_smooth(aes(x=fits, y=fits), method=lm, se=FALSE, color='red')+ #assumes mean = variance
  theme_bw()
  #Add legend
```

@@

m2<- glm.nb(data = count, nowars20~lgtrdnpi2+clqdemreg)
summary(m2)

@@

ylims <- ylim(-5, 11)  # common scale for comparison
plot_grid(autoplot(root.pois) + ylims, autoplot(root.nb) + ylims, 
          ncol = 2, labels = c("Poisson", "Neg. Binom"))

```
What does this tell us? The rootogram reflects fitted counts. If a bar doesn't reach the zero line, the model is overpredicting a certain count bin. If it exceeds the zero line, it underpredicts. So, neither of these models are perfect. BUT - the poisson model seems to both over and underpredict a lot. The negative binomial model on the other hand definitely over predicts, but not as much.

#Predicted Counts

@@

#What if we wanted a single point estimate of our predicted count of wars given our IVs?
mean(pc_sims)

#And the upper and lower boundaries of a confidence interval based on this simulation?
quantile(pc_sims, prob= 0.025) 
quantile(pc_sims, prob= 0.975) 
```

How do we predict counts over the range of a variable? Like we always have! Let's say we want to know what happens to the predicted number of counts as the average number of democratic cliques varies from its min to its max.
```{r}
temp_data <- count

@@

#we could do this in one step, but I like keeping all of the simulated counts so
#we can do whatever we want with the distribution of predicted counts
plot_data <- t(apply(pc_dem_cl, 1, quantile, probs = c(.1, .5, .9))) %>% as.data.frame()
plot_data$dem_cl <- dem_cl
```

```{r}
#Now to plot: 
colnames(plot_data)<-c("low", "med", "high", "Dem_Cliques")

#Collapse information into 3 vectors: for each value of trade in our data, 
#give us the average point estimate (expected count), the low bound of a confidence interval, and the high bound.

pc_plot <- ggplot(data=plot_data) + 
  labs(x = "Average Number of Democratic Cliques", 
       y = "Predicted Count of Intl. Wars") +
  geom_line(aes(x = Dem_Cliques, y = med, 
                color = "Number of Interstate Wars"), size = 1) +
  geom_ribbon(aes(x = Dem_Cliques, ymin = low, ymax = high, 
                  fill = "80% Confidence Interval (Simulation)"), alpha = .5) +
  theme_bw() +
  theme(legend.title = element_text(), 
        legend.position = "bottom", panel.grid.minor.x=element_blank(),
        panel.grid.minor.y = element_blank()) +
  scale_fill_manual(values = "steelblue3", name = "")+
  scale_color_manual(values = "black", name = "")
pc_plot
```

>>>>

---
title: 'Lab 10: Poisson and Negative Binomial Models: Answers'
author: 'TA: Eric Parajon'
output:
  html_document:
    df_print: paged
---

@@

Load necessary packages:
```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE, messsage = FALSE)
rm(list=ls()) 
#Load in packages
pacman::p_load(tidyverse,
               haven,#for importing data
               labelled,
               MASS,
               AER,
               foreign,
               cowplot)
```

@@

```{r, message=F, warning=F, include=F}
count <- read_dta("Maoz.dta")
```

@@

Since we're using count data what should the family argument be?

```{r}
m1 <- glm(nowars20 ~ lgtrdnpi2 + clqdemreg,
          data = count,
          family = "poisson")
summary(m1)
```
Note that this tells us "Dispersion parameter for poisson family taken to be 1."

@@

ggplot(data=over, aes(x=fits, y=V2))+
  geom_point()+
  geom_smooth(aes(x=fits, y=V2), method=lm, se=FALSE, color="blue")+ #line for mean of estimates against the variance of estimates
  geom_smooth(aes(x=fits, y=fits), method=lm, se=FALSE, color='red')+ #assumes mean = variance
  theme_bw()
```

@@

m2<- MASS::glm.nb(data = count, nowars20~lgtrdnpi2+clqdemreg)
summary(m2)

@@

ylims <- ylim(-5, 11)  # common scale for comparison

#Using cowplot to put plots together with labels
plot_grid(autoplot(root.pois) + ylims, autoplot(root.nb) + ylims, 
          ncol = 2, labels = c("Poisson", "Neg. Binom"))

```

What does this tell us? The rootogram reflects fitted counts. If a bar doesn't reach the zero line, the model is overpredicting a certain count bin. If it exceeds the zero line, it underpredicts. So, neither of these models are perfect. BUT - the poisson model seems to both over and underpredict a lot. The negative binomial model on the other hand definitely over predicts, but not as much.

# Predicted Counts

@@

#What if we wanted a single point estimate of our predicted count of wars given our IVs?
mean<-mean(pc_sims)

#And the upper and lower boundaries of a confidence interval based on this simulation?
lower<-quantile(pc_sims, prob= 0.025) 
upper<-quantile(pc_sims, prob= 0.975)

#putting it together
mapply(c,mean,lower,upper)

```

How do we predict counts over the range of a variable? Like we always have! Let's say we want to know what happens to the predicted number of counts as the average number of democratic cliques varies from its min to its max.
```{r}
#Re-naming count df to temp_data for use in this
temp_data <- count

@@

#we could do this in one step, but I like keeping all of the simulated counts so
#we can do whatever we want with the distribution of predicted counts

plot_data <- t(apply(pc_dem_cl, 1, quantile, probs = c(.1, .5, .9))) %>% as.data.frame()
plot_data$dem_cl <- dem_cl

#So this give us the average point estimate (expected count), 
#the low bound of a confidence interval, and the high bound at varying numbers of democratic cliques

```

# Exercise

To do on your own:

Below, try using ggplot to plot this information we just generated. Follow these guidelines:\
- X axis: Average number of Democratic Cliques\
- Y axis: Predicted count of Intl. Wars\
- Include a geom_line() for Dem_Cliques\
- Include a geom_ribbon() with CIs (80% confidence interval)\
- Include a legend which specifies what the line is and what the CIs are, try putting that legend on the bottom

```{r}
#Now to plot:

colnames(plot_data)<-c("low", "expected", "high", "Dem_Cliques")

pc_plot <- ggplot(data=plot_data) + 
  labs(x = "Average Number of Democratic Cliques", 
       y = "Predicted Count of Intl. Wars") +
  geom_line(aes(x = Dem_Cliques, y = expected, 
                color = "Number of Interstate Wars"), linewidth = 1) +
  geom_ribbon(aes(x = Dem_Cliques, ymin = low, ymax = high, 
                  fill = "80% Confidence Interval (Simulation)"), alpha = .5) +
  theme_bw() +
  theme(legend.title = element_text(), 
        legend.position = "bottom", panel.grid.minor.x=element_blank(),
        panel.grid.minor.y = element_blank()) +
  scale_fill_manual(values = "steelblue3", name = "")+
  scale_color_manual(values = "black", name = "")
pc_plot

```
